Looking for a Signal in the Noise: Revisiting Obesity and the Microbiome

### 2016-05-27 submission/Sze_Obesity_PeerJ_2016.Rmd
edit Figure 4 legend

![Figure 1](../results/figures/Figure_1.png)\

@@

```{r Figure_2, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE, fig.path="results/figures/", dev = c("pdf", "png")}
combinedData <- read.csv("results/tables/ROCcombinedData.csv", stringsAsFactors = FALSE)

diag <- data.frame(x=seq(1, 0), y = seq(0, 1))
diag$Study <- "Control"

rocPlot <- ggplot(data=combinedData,
                  aes(x = Sensitivity, y = Specificity))
rocPlot + geom_line(aes(group=Study, color=Study)) +
  scale_colour_brewer(palette="Set1") +
  scale_x_continuous(trans = "reverse") +
  geom_line(data=diag, aes(x=x, y=y), linetype="dashed") +
  theme(axis.line=element_line(colour="black"),
       panel.grid.major = element_blank(),
       panel.grid.minor = element_blank(),
       panel.border = element_blank(),
       panel.background = element_blank(),
       legend.key = element_rect(fill = "white"))

**Figure 2: Individual and Combined comparison of Obese and Non-Obese groups for Shannon Diversity and B/F Ratio.** A) For Shannon diversity only Baxter (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'baxter' & alpha_tests.summary$metric == 'shannon'), 'p.value'], 2), nsmall = 2)` and Goodrich (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'goodrich' & alpha_tests.summary$metric == 'shannon'), 'p.value'], 5), nsmall = 5)`) studies indivdiually were signficantly different for obese and non-obese groups.  When the data sets were combined and study effect was corrected for,  there was a significant difference between Non-Obese and Obese groups (P-value = `r format(round(alpha_composite.summary['shannon', 'composite_p'], 5), nsmall = 5)`). B) For the B/F ratio there was no individual study that was significant nor was their a difference when the data was combined and study effect corrected for between Non-Obese and Obese groups (P > 0.05).

```{r Figure_3, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE,  fig.path="results/figures/", dev = c("pdf", "png")}

grid.arrange(HRRG1, BFRRG1, ncol=2)

@@

```{r Figure_4, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE,  fig.path="results/figures/", dev = c("pdf", "png")}
# Section A
p1 <- ggplot(group.data.test, aes(obese, ZH))
p2 <- p1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
p3 <- p2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
p4 <- p3 + xlab("") + ylab("Z-score Normalized H") + scale_x_discrete(breaks=c("No", "Yes"), labels=c("Non-Obese", "Obese"))
p5 <- p4 + guides(fill=FALSE)
p6 <- p5 + theme(axis.line=element_line(colour="black"),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank())
p7 <- p6 + ggtitle('A') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section B
bf1 <- ggplot(group.data.test, aes(obese, ZLogBF))
bf2 <- bf1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
bf3 <- bf2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
bf4 <- bf3 + xlab("") + ylab("Z-score Normalized Log B/F Ratio") +
  scale_x_discrete(breaks=c("No", "Yes"), labels=c("Non-Obese", "Obese"))
bf5 <- bf4 + guides(fill=FALSE)
bf6 <- bf5 + theme(axis.line=element_line(colour="black"),
                   panel.grid.major = element_blank(),
                   panel.grid.minor = element_blank(),
                   panel.border = element_blank(),
                   panel.background = element_blank())
bf7 <- bf6 + ggtitle('B') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

grid.arrange(p7, bf7, ncol=2)
```

**Figure 4: Pooled Analysis of Shannon Diversity and B/F ratio by either Obesity Status or BMI Groups.**  **A)** Z-score Normalized Shannon Diversity and Obesity. There was a significant difference between non-obese (n = `r length(which(alpha.data.ALL$obese == FALSE))`) and obese groups (n = `r length(which(alpha.data.ALL$obese == TRUE))`) (P-value = `r format(round(alpha_composite.summary['shannon', 'composite_p'], 4), nsmall=4)`). **B)** Z-score Normalized B/F Ratio and Obesity. There was no significant difference between the non-obese (n = `r length(which(alpha.data.ALL$obese == FALSE))`) and obese (n = `r length(which(alpha.data.ALL$obese == TRUE))`) group based on B/F ratio (P-value = `r format(round(alpha_composite.summary['bf_ratio', 'composite_p'], 2), nsmall=2)`).

@@

**Figure 5. Summary of Power and Sample Size Simulations for Shannon Diversity for Non-obese versus Obese**  P-values listed in the legend represent the outcome of a wilcoxson rank sum test between non-obese and obese for each specific data set.  **A)** The dotted lines represent the actual effect size for each specific study.  **B)** The dotted lines represent the needed n to achieve an 80% power with the actual study effect size.

>>>>

@@

**Figure 2: Individual and Combined comparison of Obese and Non-Obese groups for Shannon Diversity and B/F Ratio.** **A)** For Shannon diversity only Baxter (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'baxter' & alpha_tests.summary$metric == 'shannon'), 'p.value'], 2), nsmall = 2)` and Goodrich (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'goodrich' & alpha_tests.summary$metric == 'shannon'), 'p.value'], 5), nsmall = 5)`) studies indivdiually were signficantly different for obese and non-obese groups.  When the data sets were combined and study effect was corrected for,  there was a significant difference between Non-Obese and Obese groups (P-value = `r format(round(alpha_composite.summary['shannon', 'composite_p'], 5), nsmall = 5)`). **B)** For the B/F ratio there was no individual study that was significant nor was their a difference when the data was combined and study effect corrected for between Non-Obese and Obese groups (P > 0.05).

@@

**Figure 4: ROC Curves for Each Study based on Classification of Non-Obese or Obese Groups at the Genus or OTU level.**  **A)** No significant difference was found between the different ROC curves based on Study for the genus level.  The overall AUCs ranged from `r round(min(rand_forest.genus.summary$auc), 2)` to `r round(max(rand_forest.genus.summary$auc), 2)` with a mean of `r round(mean(rand_forest.genus.summary$auc), 2)`. **B)** Similarly no difference was found between the ROC curves based on Study at the OTU level.  The overall mean of the AUCs were slightly higher overall than the genus level `r round(mean(rand_forest.otu.summary$auc), 2)` but the range was greater going from `r round(min(rand_forest.otu.summary$auc), 2)` to `r round(max(rand_forest.otu.summary$auc), 2)`.

@@

**Figure 5: Summary of Power and Sample Size Simulations for Shannon Diversity for Non-obese versus Obese**  P-values listed in the legend represent the outcome of a wilcoxson rank sum test between non-obese and obese for each specific data set.  **A)** The dotted lines represent the actual effect size for each specific study.  **B)** The dotted lines represent the needed n to achieve an 80% power with the actual study effect size.
